---
title: "Update LCM abstracts"
author: "Lambda Moses"
date: "3/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I used to update LCM abstract data by querying PubMed and downloading everything again and geocoding everything again. That's not very efficient, because I only need to query abstracts and geocode for new entries. So here I'll do that and later I'll add the code to the museumst package.

```{r}
library(tidyverse)
library(XML)
library(ggmap)
library(easyPubMed)
library(lubridate)
library(museumst)
```

```{r}
my_query <- "((laser capture microdissection) OR (laser microdissection)) AND ((microarray) OR (transcriptome) OR (RNA-seq))"
my_ids <- get_pubmed_ids(my_query, api_key = Sys.getenv("PUBMED_API_KEY"))
```

```{r}
my_xmls <- fetch_pubmed_data(my_ids, format = "xml", retmax = 2000)
xmlx_list <- articles_to_list(my_xmls)
```

```{r}
con <- file("pubmed_lcm.xml")
writeLines(my_xmls, con = con)
close(con)
```

```{r}
df <- map_dfr(xmlx_list, article_to_df, max_chars = 0)
```

```{r}
saveRDS(df, "lcm_info.rds")
```

For now, I only care about the address of the first author.
```{r}
df_1st <- df %>% 
  group_by(pmid, title) %>% 
  mutate(author_ord = seq_along(title)) %>% 
  filter(author_ord == 1) %>% 
  select(-author_ord)
```

Also, I want to exclude reviews, which means I need to get article types.
```{r}
xmls <- xmlParse("pubmed_lcm.xml")
types <- xpathApply(xmls, "//PublicationTypeList")
types <- lapply(types, function(x) xmlSApply(x, xmlValue))
is_review <- map_lgl(types, ~ any(.x == "Review"))
```

What proportion is reviews?
```{r}
mean(is_review)
```

```{r}
# Old
data("lcm_abstracts")
```

I'll only keep entries not already in the collection and are not reviews.
```{r}
inds_keep <- !is_review & (!df_1st$pmid %in% lcm_abstracts$pmid)
```

```{r}
df_1st <- df_1st[inds_keep, ]
```

```{r}
df_1st <- df_1st %>% 
  unite(col = "date_published", year, month, day, sep = "-") %>% 
  mutate(date_published = ymd(date_published))
```

```{r}
# Old
data("lcm_city_gc")
```

Just learnt this from Stack Overflow.
```{r}
df_gc <- df_1st %>% 
  mutate(geodata = purrr::map(address, 
                              ~ {if (is.na(.x)) NA else geocode(.x, output = "all")})) %>% 
  filter(!is.na(address))
```

```{r}
df_gc <- df_gc %>% 
  mutate(address_components = purrr::map(geodata, list("results", 1, "address_components")),
         to_rm = map_lgl(address_components, is.null)) %>% 
  filter(!to_rm) %>% 
  select(-to_rm)
```

```{r}
df_gc <- df_gc %>% 
  mutate(address_components = purrr::map(address_components, 
                                  ~ as_tibble(transpose(.x)) %>% 
                                    unnest(c("long_name", "short_name")) %>% 
                                    unnest("types") %>% 
                                    unnest("types")),
         country = map_chr(address_components, ~ .x$long_name[.x$types == "country"]),
         `state/province` = map_chr(address_components, 
                         ~ {out <- .x$long_name[.x$types == "administrative_area_level_1"]
                         if (length(out) == 0) NA_character_ else out
                         }))
df_gc <- df_gc %>% 
  mutate(city = map_chr(address_components, 
                        ~ {out <- .x$long_name[.x$types == "postal_town"]
                        if (length(out) == 0)
                          out <- .x$long_name[.x$types == "locality"]
                        if (length(out) == 0) 
                          out <- .x$long_name[.x$types == "administrative_area_level_2"]
                        if (length(out) == 0)
                          out <- .x$long_name[.x$types == "administrative_area_level_1"]
                        if (length(out) == 0) out <- NA_character_
                        out
                        }))
```

Sounds like really dumb, but I'm going to geocode the cities again. 

```{r}
df_gc <- df_gc %>% 
  mutate(`state/province` = str_remove(`state/province`, "\\s(Sheng)|(Shi)$"),
         city = str_remove(city, "\\s(City)|(Shi)$")) %>% 
  ungroup()
```

```{r}
# Only do cities not already in the collection
cities <- df_gc %>%
  select(city, `state/province`, country) %>%
  distinct() %>%
  filter(!is.na(city)) %>%
  unite(col = "city2", city, `state/province`, country, remove = FALSE, sep = ", ") %>%
  mutate(city2 = str_remove(city2, "NA, ")) %>% 
  anti_join(lcm_city_gc, by = "city2")
cities_gc <- geocode(cities$city2)
cities_gc <- cbind(cities, cities_gc)
cities_gc <- sf::st_as_sf(cities_gc, coords = c("lon", "lat"), crs = sf::st_crs(4326)) %>%
  select(city, city2, `state/province`, country, geometry)
```

```{r}
cities_gc <- bind_rows(cities_gc, lcm_city_gc)
```

```{r}
saveRDS(cities_gc, "lcm_city_gc.rds")
```

I'll only get the abstracts of new entries
```{r}
abstracts <- map_dfr(xmlx_list[inds_keep], article_to_df, getAuthors = FALSE, 
                     max_chars = 1e6)
```

```{r}
abstracts <- abstracts %>% 
  select(pmid:journal)
```

```{r}
abstracts <- abstracts %>% 
  unite(col = "date_published", year, month, day, sep = "-") %>% 
  mutate(date_published = ymd(date_published)) %>% 
  inner_join(df_gc[, c("pmid", "address", "country", "state/province", "city")])
```

Also remove empty abstracts
```{r}
abstracts <- abstracts %>% 
  filter(abstract != "")
```

```{r}
abstracts <- abstracts %>% 
  mutate(date_num = as.numeric(date_published))
```

```{r}
abstracts <- abstracts %>% 
  unite(col = "city2", city, country, sep = ", ", remove = FALSE) 
```

```{r}
# append to old collection
abstracts <- bind_rows(abstracts, lcm_abstracts)
```

```{r}
lcm_abstracts <- abstracts
lcm_city_gc <- cities_gc
```

```{r}
usethis::use_data(lcm_abstracts, overwrite = TRUE)
usethis::use_data(lcm_city_gc, overwrite = TRUE)
```

