---
title: "Prequel Analysis"
author: "Lambda Moses"
date: "5/25/2020"
output: html_document
params:
  sheet: "Prequel analysis"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here I explore the history and geography of spatial transcriptomics based on microdissection through the metadata I collected.
```{r}
library(tidyverse)
library(googlesheets4)
library(lubridate)
library(magick)
library(scales)
library(zeallot)
source("code/metadata_plotting.R")
theme_set(theme_bw())
```

# Import data
```{r}
gs4_deauth()
sheet <- read_sheet("https://docs.google.com/spreadsheets/d/1sJDb9B7AtYmfKv4-m8XR7uc3XXw_k4kGSout8cqZ8bY/edit#gid=566523154", sheet = params$sheet)
```

Add the year column as this will be used for plotting rather than the full date.
```{r}
sheet <- sheet %>% 
  mutate(year = year(date_published))
```

# Number of publications
How many publications per year?
```{r}
# Remove duplicates as different datasets in the same publication have their own rows
publications <- sheet %>% 
  select(year, title, journal, repo:department) %>% 
  distinct()
```

```{r}
anyDuplicated(publications$title)
```

How many publications are there in total in this sheet?
```{r}
nrow(publications)
```

```{r}
# For maps later in this notebook
c(inst_gc, city_gc) %<-% geocode_inst_city(publications)
pubs_on_map2 <- partial(pubs_on_map, inst_gc = inst_gc, city_gc = city_gc)
```

## Overall

```{r}
pubs_per_year(publications)
```

Apparently, the golden age of prequel data analysis is gone.

## By category
By the category of problem to address
```{r}
category <- sheet %>% 
  mutate(category = str_split(category, pattern = "(;|,)\\s")) %>% 
  select(year, title, journal, category, country, institution) %>% 
  unnest(cols = "category")
```

```{r}
pubs_per_cat(category, "category")
```

```{r, fig.width=4, fig.height=8}
pubs_per_year(category, facet_by = "category")
```

## By species
These are the species the computational methods were demonstrated on
```{r}
species_img <- readRDS("output/species_img.rds")
```

```{r}
species <- sheet %>% 
  select(year, title, journal, species, country, institution) %>% 
  filter(!is.na(species)) %>% 
  mutate(species = str_split(species, "; ")) %>% 
  unnest(cols = "species") %>%
  distinct()
```

```{r}
unique(species$species)
```

```{r}
pubs_per_cat(species, category = "species")
```

```{r}
top_species <- species %>% 
  count(species) %>%
  top_n(5, n) %>% 
  arrange(desc(n)) %>% 
  pull(species)
species_img$image <- purrr::map(species_img$image_paths, image_read)
```

```{r}
species %>% 
  filter(species %in% top_species) %>% 
  left_join(species_img, by = "species") %>% 
  pubs_per_cat(category = "species", isotype = TRUE, img_unit = 5)
```

I know, for almost all cases, it's either for the BDGP atlas or the Allen Brain Atlas.

```{r}
pubs_per_year(species, facet_by = "species")
```

I don't think it's significant, but it looks like the mouse papers came slightly later than Drosophila papers.

## By journal
```{r}
sort(table(publications$journal), decreasing = TRUE)
```

## Location
### General
Just some barplots for number of publications per institution, city, and country. 
```{r}
pubs_per_cat(publications, "country")
```

What, Sweden isn't here? How about per capita? Actually this probably does not say much since many researchers in the West are immigrants. Then it raises the burning political question of who counts as the "population".

```{r}
pubs_per_capita(publications, plot = "bar")
```

How about country over time? There might not be enough publications to show a trend.
```{r, fig.height=8, fig.width=4}
pubs_per_year(publications, facet_by = "country")
```
Now look at cities. 

```{r}
pubs_per_cat(publications, "city")
```

Institutions
```{r}
pubs_per_cat(publications, "institution")
```

It's very different from the institutions for the original trilogy analysis.

### Worldwide
```{r}
pubs_on_map2(publications)
```

Let me also plot the per capita thing on a map, as a choropleth
```{r}
pubs_per_capita(publications)
```

Break down by species
```{r, fig.width=4, fig.height=8}
pubs_on_map2(species, facet_by = "species", ncol = 1)
```


Probably this does not tell us much since there's only 1 study for some species. Again, we see the received wisdom that research is mostly confined to the West, especially Western Europe, west coast of the US, and New England, but not as much in the other parts of the "West".

### Europe
Plot a map just for Europe. A problem here is that if I simply filter the `world` sf object, I'll get some islands away from what we usually think is Europe. Kind of feel bad for Russians since Russia is so large that it makes the map look bad and I don't have a paper from Russia in this spreadsheet. 

```{r}
pubs_on_map2(publications, zoom = "europe")
```

```{r}
pubs_per_capita(publications, "europe")
```

```{r}
pubs_on_map2(species, zoom = "europe", facet_by = "species")
```

Maybe I just made this plot for fun. Not sure what to say about it.

### USA
Also a plot just for America. I did not encounter a publication in this spreadsheet that is from Hawaii or Alaska, but I'll include Hawaii and Alaska in the map anyway in case I get one in the future.

```{r}
pubs_on_map2(publications, zoom = "usa")
```

```{r}
pubs_per_capita(publications, "usa")
```

```{r}
pubs_on_map2(species, zoom = "usa", facet_by = "species")
```

# Word cloud
## Titles
```{r}
plot_wordcloud(sheet)
```

## Summaries
```{r}
plot_wordcloud(sheet, col_use = "summary")
```

```{r}
plot_wordcloud(sheet, col_use = "core_principle")
```

## Tissues
```{r}
plot_wordcloud(sheet, col_use = "tissue")
```

## By species
```{r}
plot_wordcloud(sheet, species_use = "Drosophila melanogaster")
```

```{r}
plot_wordcloud(sheet, species_use = "Mus musculus")
```

## Over time

```{r}
range(sheet$date_published)
```

### Titles
```{r}
plot_wordcloud(sheet, year_min = 2001, year_max = 2010)
```

```{r}
plot_wordcloud(sheet, year_min = 2010, year_max = 2015)
```

```{r}
plot_wordcloud(sheet, year_min = 2015, year_max = 2021)
```

### Summaries
```{r}
plot_wordcloud(sheet, col_use = "summary", year_min = 2001, year_max = 2010)
```

```{r}
plot_wordcloud(sheet, col_use = "summary", year_min = 2010, year_max = 2015)
```

```{r}
plot_wordcloud(sheet, col_use = "summary", year_min = 2015, year_max = 2021)
```

## Department names
```{r}
institution_words <- readRDS("output/inst_words.rds")
```

```{r}
plot_wordcloud(sheet, col_use = "department", other_stop_words = institution_words)
```
